\textbf{\Large Computer Vision-based System for Impaired Human Vision Compensation}

### 2018-10-14 bare_conf.tex
text updates in Abstract, Results and Discussion

\begin{abstract}
%\boldmath
In this short article we describe and discuss high level architecture of computer vision-based system for partial compensation of lost or impaired human vision. Combining standard smartphone device, external deep learning based vision processor and audio-tactile user interface into compact visual perception compensation device the suggested architecture may be advantageous both for researchers and industrial product developers.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\begin{IEEEkeywords}
Computer vision; Deep learning; Mobile application; Aid for blind and visually impaired; Audio feedback; Tactile feedback; Impaired human vision compensation.%
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}

More than $250$ million people worldwide have moderate to severe vision impairment, while ($\approx 36$ million are blind) \cite{Bourne}. During the past decades significant effort was devoted to develop computer vision and other sensor based aids (e.g. \cite{Caraiman}, \cite{Csapo}, \cite{Poggi}, \cite{Zientara}) for helping the blind and visually impaired users to better perceive the surrounding world. However, computer vision is a rapidly evolving field; systems based on the aforementioned approaches become outdated in a relitively short time and often lack accuracy and reliability in real-world conditions in comparisson to the state-of-the-art technologies. In this article we describe a system, which employs modern computer vision techniques for compensation of lost or impaired vision function in humans. Many of the previously proposed electronic aids for the blind count on highly specialized hardware, for instance smart glasses \cite{Zientara}, Microsoft Kinnect sensors \cite{Owayjan}, helmet-mounted photo sensors and cameras \cite{Dunai}. Such a systems often lack convenience and bring additional complexity into daily tasks, which visually impaired people currently manage by the help of a white cane. This project puts major emphasis on developing a tool, which seamlessly integrates with the hardware visually impaired persons already have on-hand and is used to (smartphone). This paper describes a high level architecture of the technology aid for visually impaired people utilising the latest achievemets in the field of computer vision.

@@

Following the principles of the participatory design \cite{Kensing}, \cite{Carroll} we included the end-users into the design process of the system. Two visually impaired persons participated in discussion and focus group meetings together with the project team to identify the key properties of the solution. During the six discussion sessions major attention was paid to:
\begin{enumerate}
\item Functionality and key features of the solution addressing real-life scenarios of visually impaired persons;
\item Technical feasibility, connecting user needs to the latest development in computer vision and assisted technologies;
\item Interface between the visually impaired person and the assisted technology;
\item Appearance, usability and potential costs of the final product. 
\end{enumerate}
Design process involved testing of the existing mobile phone app (Seeing AI) for assisting visually impaired persons. Semi-structured tests were performed by the visually impaired participants accompanied by a researcher from a project group. Strengths and weaknesses of the existing solutions are reflected in the design of the proposed system.

@@

% Man atrodo, kad cia du esminiai sistemos komponentai, kuriuos reiktu akcentuoti:
The cornerstones of the proposed solution are its user interface and vision processing system, consisting of external server, equipped with a set of computer vision algorithms, selected and combined according to the above methodology.

% User interface
User interface consists of a wearable forehead belt, integrating RGB and depth cameras, inertial measurement unit (IMU), tactile feedback device and bone conduction headphones. Although more expensive than standard ones, bone conduction headphones do not block the ears and allow the user to perceive audio information from the surrounding environment, which is essential for the visually impaired people. The tactile feedback in certain cases may improve the quality of users perception (e.g. by indicating the direction of detected object with vibromotor in the corresponding area on the forehead belt, instead of providing longer and less convenient to perceive audio description of the event). UI can interface smartphone component via USB or Wifi connection.

% External vision processor
%visus sutrumpinimus reikia pirma parasyt pilnai, pvz Neural Network (NN)
Although modern high-end smartphones are quite powerful computers themselves, many computer vision algorithms often demand hardware of significantly higher computational capacity (e.g. server with graphics processing unit (GPU)). This motivates us to employ an external server for vision data processing. In our design we assume such a server with a set of computer vision algorithms implemented and exposed via web services: faster regions with convolutional neural network features (faster R-CNN) object detector \cite{Ren}, trained to detect important objects, convolutional and recurrent neural network (CNN and RNN)-based scene description \cite{Liu}, which provides textual annotation of a given RGB image, place recognition \cite{Ohn-Bar}, face detection and recognition \cite{Amos}, obstacle detection \cite{Laina}, action recognition \cite{Valor} among others. After completing the analysis it's results are transmitted back to the smartphone component and represented to the user through the aforementioned interface.

% How user controls system
Gesture or voice command recognition can be used to select system's operation regime (e.g. optical character recognition (OCR), currency recognition, object detection, navigation aid).

% Potential applications
Social networking functionality was accepted as a potentially important feature allowing other users/volunteers to describe certain objects (for instance, temporary road works) with exact location and description. Navigating around such obstacles can be challenging for computer vision-based tools and may decrease user's trust in the tool. Integration of additional information from other users, governmental organizations and volunteers could be of high value for the visually impaired people.

@@

Discussions on functionality and key features of the solution revealed a gap between the aims of the existing tools, which are often focused on advanced navigation and scene description features, and actual user needs. For example, the end users highlighted that a simplistic solution for navigating is lacking. 
%They were not excited about advanced object detection and scene description techniques, which may overcomplicate their daily life. 
Major focus was put on a easy to use and minimalistic tool, which can be trusted. Reliable object detection, direction and approximate distance to an obstacle were highlighted as major requirements. To our surprise, only 5-8 distinctive object classes (for instance, buss stop, pedestrian crossing, doors, stairs, etc) were of major importance while navigating. Additional information, such as type of a passing car, blossoming flowers in the park or a person riding a bicycle was considered as overwhelming and distractive.

%Social networking functionality was accepted as a potentially important feature allowing other users/volunteers to describe certain objects (for instance, temporary road works) with exact location and description. Navigating around such obstacles can be challenging for computer vision-based tools and may decrease user's trust in the tool. Integration of additional information from other users, governmental organizations and volunteers could be of high value for the visually impaired people.

%Povilai, paziurek ar cia viskas OK, pridek ka nors nuo saves
% Pamineti mobilenet ir pan.
Computer vision technology is under rapid development and during the last years made a major breakthrough in performance and efficiency. Deep learning neural networks, which are de facto standard in modern image/video processing, are now well supported by software libraries and dedicated processing hardware is built-in even in relatively low-power computational devices, such as smartphones. The application of both open source and commercial computer vision software libraries (e.g. \cite{Verilook} face recognition kit) could reduce development costs. These devices are also equipped with 4G Internet connectivity and sufficient computational power to perform initial data processing locally. The maturity of the aforementioned technologies suggest that our proposed computational aid for visually impaired persons may be highly feasible.

@@

Appearance, usability and cost of the end product are of major importance to the users. The proposed assistive technology should supplement the tool that visually impaired persons used for decades - the white cane. Using a white cane requires little to none training, has very low costs and is relatively reliable. It provides information on the surrounding objects 1-2 meters in front of the user. While the cane can used to detect a nearby obstacle, electronic travel aid could be beneficial for longer range (2-10m) route planning and object detection. However, it should integrate seamlessly into the existing navigation practices of visually impaired people.

@@

In this article we outlined an idea of complex computer vision-based  sensor architecture, which may help to partially compensate impaired or lost human sight. Main advantages of therein suggested system are: ability to use efficient (but still computationally intensive) modern computer vision algorithms via the Internet connection, present the output of video processing to the user through a combined audio/tactile interface. Comparing to an existing commercial products e.g. \cite{orcam}, \cite{horus}, our approach is advantageous, since it does not rely only on computational capacity of mobile handset. Such a combination of mobile interface and remote vision processor will be even more efficient in the future, with switching of mobile networks to 5G. Combined audio/tactile interface also may be more convenient for the users, since it allows to provide the feedback more efficiently, comparing to purely audio or purely tactile means.

The suggested system does not aim to replace the main travelling aid of visually impaired people - the white cane. Instead, this computer vision tool aims to enhance and push the perception of the surounding environment boundary from 1-2 meters (achieved by using the white cane) to 2-10 meters. It may improve route planning and identification of objects of interest (for instance, doors, stairs, elevators, bus stopps, etc.). This functionality corresponds with the main requirement for electronic travelling aids identified by the end users - direction and distance estimation to a selected object. The proposed hardware architecture can be used as a basis for various computer vision-based software modules, aiming to assist visually impaired users in daily activities (e.g. outdoor/indoor navigation, object/face recognition, obstacle detection, etc.).

>>>>

\begin{abstract}
%\boldmath
This paper presents a high level architecture for computer vision-based system for partial compensation of lost or impaired human vision. It combines standard smartphone device, external deep learning-based live video processoing infrastructure and audio/tactile user interface. The proposed architecture is based on  input from user-centered design process, involving end users into system development process. The paper discusses user needs and expectations towards electronic travelling aids for the blind and and highlights limitations of the existing solutions. The suggested architecture may be used as a basis for developing computer vision-based tools for visually impaired individuals. 
%In this short article we describe and discuss high level architecture of computer vision-based system for partial compensation of lost or impaired human vision. Combining standard smartphone device, external deep learning-based live video processoing infrastructure and audio/tactile user interface into compact visual perception compensation device .
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\begin{IEEEkeywords}
Computer vision; Deep learning; Mobile application; Aid for blind and visually impaired; Audio feedback; Tactile feedback; Impaired human vision compensation; User-centered design.%
\end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:introduction}
More than $250$ million people worldwide have moderate to severe vision impairment, while ($\approx 36$ million are blind) \cite{Bourne}. During the past decades significant effort was devoted to develop computer vision and other sensor based aids (e.g. \cite{Caraiman}, \cite{Csapo}, \cite{Poggi}, \cite{Zientara}) for helping the blind and visually impaired users to perceive the surrounding world better. However, computer vision is a rapidly evolving field; systems based on the aforementioned approaches become outdated in a relitively short time and often lack accuracy and reliability in real-world conditions in comparisson to the state-of-the-art technologies. In this article we describe a system, which employs modern computer vision techniques for compensation of lost or impaired vision function in humans. Many of the previously proposed electronic aids for the blind count on highly specialized hardware, for instance smart glasses \cite{Zientara}, Microsoft Kinnect sensors \cite{Owayjan}, helmet-mounted photo sensors and cameras \cite{Dunai}. Such systems often lack convenience and bring additional complexity into daily tasks, which visually impaired people currently manage by the help of a white cane. This project puts major emphasis on developing a tool, which seamlessly integrates with the hardware visually impaired persons already have on-hand and is used to (smartphone). This paper describes a high level architecture of the technology aid for visually impaired people utilising the latest achievemets in the field of computer vision.

@@

Following the principles of the participatory design \cite{Kensing}, \cite{Carroll} we included the end-users into the design process of the system. Two visually impaired persons participated in discussion and focus group meetings together with the project team to identify the key properties of the solution. During the six discussion sessions major attention was paid to:
\begin{enumerate}
\item Functionality and key features of the solution addressing real-life scenarios of visually impaired persons;
\item Technical feasibility, connecting user needs to the latest development in computer vision and assisted technologies;
\item Interface between the visually impaired person and the assisted technology;
\item Appearance, usability and potential costs of the proposed product. 
\end{enumerate}
Design process included testing of the existing mobile phone app (Seeing AI) for assisting visually impaired people. Semi-structured tests were performed by the visually impaired participants accompanied by a researcher from the project group. Strengths and weaknesses of the existing solutions are reflected in the design of the proposed system.

@@

% Man atrodo, kad cia du esminiai sistemos komponentai, kuriuos reiktu akcentuoti:
%The cornerstones of the proposed solution are its user interface and vision processing system, consisting of external server, equipped with a set of computer vision algorithms, selected and combined according to the above methodology.

% User interface
UI consists of a wearable forehead belt, integrating RGB and depth cameras, inertial measurement unit (IMU), tactile feedback device and bone conduction headphones. Although more expensive than standard ones, bone conduction headphones do not block the ears and allow the user to perceive audio information from the surrounding environment and the system at the same time, which is essential for the visually impaired people. Some types of guidance information may be more efficiently perceived through tactile feedback device than audio channel. For instance, indicating the direction of a detected object with a vibromotor on the corresponding area on the forehead belt, instead of providing audio description. More research is needed to clarity how variuos types of guidance information should be represented to the user in the most convenient and efficient way.

%The tactile feedback may improve the quality of users perception (e.g. by ). UI can interface smartphone component via USB or Wifi connection.

% External vision processor
%visus sutrumpinimus reikia pirma parasyt pilnai, pvz Neural Network (NN)
Although modern high-end smartphones are powerful computational devices, many computer vision algorithms  require hardware of significantly higher processing power (e.g. a server equipped with a high-end graphics processing unit (GPU)). Therefore, an external server is used for live vision data processing. In our design we envision a computational server hosting set of computer vision algorithms implemented and exposed via web services: faster regions with convolutional neural network features (faster R-CNN) object detector \cite{Ren}, trained to detect important objects, convolutional and recurrent neural network-based (CNN and RNN) scene descriptor \cite{Liu}, which provides textual annotation of a given RGB image, place recognition \cite{Ohn-Bar}, face detection and recognition \cite{Amos}, obstacle detection \cite{Laina}, action recognition \cite{Valor}, among others. After video analysis is completed on the server side, results are transmitted back to the smartphone component and represented to the user through the aforementioned UI.

% How user controls system
The user interacts with the system through gestures, voice commands or a specifically designed user interface on the smartphone. User input is required for selecting the operation mode of the system, i.e. swithching between optical character recognition (OCR), currency recognition, object detection, navigation aid and other available functions.

% Potential applications
Social networking functionality was highlighted as a potentially important feature allowing other users/volunteers contribute to up-to-date and high quality guidance (for instance, temporary road works) with exact location and description. Navigating around such obstacles can be challenging for computer vision-based tools and may decrease user's trust in the tool. Integration of additional information from other users, governmental organizations and volunteers could be of high value for the visually impaired people.

@@

Discussions on the key features of the proposed solution revealed a gap between the functionality of the existing tools, which are often focused on advanced navigation and scene description features, and actual user needs. For example, the end-users highlighted that a simplistic solution for navigating is lacking. Major focus was put on an easy-to-use and minimalistic tool, which can be trusted. Reliable object detection, direction and approximate distance to an obstacle were highlighted as major requirements. To our surprise, only 5-8 distinctive object classes (for instance, buss stop, pedestrian crossing, doors, stairs, etc) were of major importance while navigating. Additional information, such as type of a passing car, blossoming flowers in the park or a person riding a bicycle was considered as overwhelming and distractive.

%Povilai, paziurek ar cia viskas OK, pridek ka nors nuo saves
% Pamineti mobilenet ir pan.
Computer vision technology is under rapid development and during the last years made a major breakthrough in performance and efficiency. Deep learning neural networks, which are de facto standard in modern image/video processing, are now well supported by software libraries and dedicated processing hardware is built-in even in relatively low-power computational devices, such as smartphones. 
%Andrius: as sita sakini ismesciau, man jis nelabai cia tinka
The application of both open source and commercial computer vision software libraries (e.g. \cite{Verilook} face recognition kit) could reduce development costs. 
%Andrius: site sakiniai turetu eiti iskart po "... such as snartphones."
These devices are also equipped with 4G Internet connectivity and sufficient computational power to perform initial data processing locally. The maturity of the aforementioned technologies suggest that our proposed computational aid for visually impaired persons may be highly feasible.

@@

Appearance, usability and cost of the end product are of major importance to the users. The proposed assistive technology should supplement the tool that visually impaired persons used for decades - the white cane. Using a white cane requires little to no training, has very low costs and is relatively reliable. It provides information on the surrounding objects 1-2 meters in front of the user. While the cane can be used to detect nearby obstacles, electronic travel aid could be beneficial for longer range (2-10m) route planning and object detection. However, it should integrate seamlessly into the existing navigation practices of visually impaired people.

@@

%In this article we outlined an idea of complex computer vision-based sensor architecture, which may help to partially compensate impaired or lost human sight.

This paper outlines a high level architecture of a computer vision-based system, which may help to partially compensate impaired or lost human vision. The main advantages of therein suggested system are: ability to use efficient (but still computationally intensive) modern computer vision algorithms via the Internet connection and present the output of the video processing to the user through a combined audio/tactile interface.

Similar functionality has been previuosly addressed by research communities \cite{Caraiman}, \cite{Csapo}, \cite{Poggi}, \cite{Zientara} and industy \cite{orcam}, \cite{horus}. The main difference between the aforementioned commercial solutions and the proposed system is the ability to utilize external resources for computationally intensive live video processing tasks. While the availability of the computational power could be seen as the main advantage, it comes with high cost - dependency on a well-functioning mobile broadband. The development of the high-speed mobile networks (4G and 5G in the nearest future) is likely to make the major limitation of the proposed architecture obsolete, especially in densely populated areas containing many hazards for the blind. Combined audio/tactile interface may also be more convenient for the users, allowing to provide feedback in a more efficient way than using tactile or audio interfaces separately.

%our approach is advantageous, since it does not rely only on computational capacity of mobile handset. Such a combination of mobile interface and remote vision processor will be even more efficient in the future, with switching of the mobile networks to 5G.

It is important to emphasize that the suggested system does not aim to replace the main travelling aid of visually impaired people - the white cane. Instead, this computer vision tool aims to enhance and push the perception of the surounding environment boundary from 1-2 meters (achieved by using the white cane) to 2-10 meters. It may improve route planning and identification of objects of interest (for instance, doors, stairs, elevators, bus stopps, etc.). This functionality corresponds with the main requirement for electronic travelling aids highlighted by the end users - direction and distance estimation to a selected object. The proposed hardware architecture can be used as a basis for various computer vision-based software modules, aiming to assist visually impaired users in daily activities (e.g. outdoor/indoor navigation, object/face recognition, obstacle detection, etc.).
